---
title: "SEA rsfcMRI - Compare Variance Explained"
author: "John Flournoy"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{css, echo=FALSE}
@import url('https://fonts.googleapis.com/css?family=Didact+Gothic&display=swap');
@import url('https://fonts.googleapis.com/css?family=Fira+Code&display=swap');
body{
  font-family: 'Didact Gothic',  sans-serif;
}
pre code {
  font-family: 'Fira Code',  monospace;
}
```

```{r setup, include=FALSE}
fslong <- FALSE ### CHANGE FILE NAME TOO!
library(ggplot2)
library(data.table)
library(dplyr)
library(tidyr)
library(showtext)
library(lme4)
library(DT)
#{.tabset}

font_add_google("Didact Gothic", "Didact Gothic")
showtext_auto()

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
#https://www.instagram.com/p/CC3EvmLgt2m/?utm_source=ig_web_copy_link
apal <- paste0('#',c('2C2B2B', 'F9773B', 'FFEA8C', '1389E6', 'D2E5E7'))
jftheme <- theme_minimal() +  
    theme(text = element_text(family = 'Didact Gothic', size = 14),
          panel.background = element_rect(fill = apal[[5]], size = 0, color = apal[[2]]),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          strip.background = element_rect(fill = apal[[3]], size = 0),
          strip.text = element_text(color = '#222222'),
          axis.text =  element_text(color = apal[[1]]), axis.title = element_text(color = apal[[1]]))

get_within_net_connections <- function(d){
  sea_schaefer400_7_withinnet <- d
  
  sub_regex <- '[LR]H_[A-Za-z]+_*(.*)'
  net_regex <- '[LR]H_([A-Za-z]+)_*(?:.*)'
  hem_regex <- '([LR]H)_[A-Za-z]+_*(?:.*)'
  sea_schaefer400_7_withinnet_nets <- distinct(sea_schaefer400_7_withinnet, network1)
  sea_schaefer400_7_withinnet_nets[, sub1 := gsub(sub_regex, '\\1', network1)]
  sea_schaefer400_7_withinnet_nets[, net1 := gsub(net_regex, '\\1', network1)]
  sea_schaefer400_7_withinnet_nets[, hem1 := gsub(hem_regex, '\\1', network1)]
  setkey(sea_schaefer400_7_withinnet, network1)
  setkey(sea_schaefer400_7_withinnet_nets, network1)
  sea_schaefer400_7_withinnet <- sea_schaefer400_7_withinnet_nets[sea_schaefer400_7_withinnet]
  
  setnames(sea_schaefer400_7_withinnet_nets, c('network2', 'sub2', 'net2', 'hem2'))
  setkey(sea_schaefer400_7_withinnet, network2)
  setkey(sea_schaefer400_7_withinnet_nets, network2)
  sea_schaefer400_7_withinnet <- sea_schaefer400_7_withinnet_nets[sea_schaefer400_7_withinnet]
  
  sea_schaefer400_7_withinnet <- sea_schaefer400_7_withinnet[net1 == net2 & hem1 == hem2]
  sea_schaefer400_7_withinnet[, c('network1', 'network2') := NULL]
  sea_schaefer400_7_withinnet[, z := atanh(r)]
  sea_schaefer400_7_withinnet[, H := dplyr::case_when(hem1 == 'RH' ~ -1,
                                                      hem1 == 'LH' ~ 1)]
  if(!dim(distinct(sea_schaefer400_7_withinnet, net1, net2, hem1, hem2))[[1]] == 14){
    stop('Incorrect number of networks')
  }
  return(sea_schaefer400_7_withinnet)
}

get_variance_table <- function(model_fit){
  vc <- VarCorr(model_fit)
  id_v <- vc$id['(Intercept)','(Intercept)']
  idsess_v <- vc$`sess:id`['(Intercept)','(Intercept)']
  s_2 <- sigma(model_fit)^2
  total_RE <- sum(unlist(lapply(vc, diag)))
  other_RE <- total_RE - id_v - idsess_v
  total <- total_RE + s_2
  rez <- data.frame(source = rep(c('ID', 'ID/Session', 'Other RE', 'Residual'),2),
                    out_of = factor(c(rep('total', 4), rep('RE', 4)), levels = c('total', 'RE')),
                    variance = c(id_v, idsess_v, other_RE, s_2, id_v, idsess_v, other_RE, NA),
                    sd = c(id_v, idsess_v, other_RE, s_2, id_v, idsess_v, other_RE, NA)^.5,
                    percent = c(c(id_v, idsess_v, other_RE, s_2)*100 / total,
                                c(id_v, idsess_v, other_RE, NA)*100 / total_RE), 
                    stringsAsFactors = FALSE)
  if(other_RE == 0){
    rez <- rez[rez$source != 'Other RE',]
  }
  return(rez)
}

dt_options <- list(rownames = FALSE,
                   filter = 'top',
                   class = 'cell-border stripe',
                   extensions = 'Buttons', 
                   options = list(dom = 'Bfrtip', buttons = c('csv')))

dt_table <- function(data, caption, dt_options = NULL){
  options_list <- c(dt_options[['options']], list(dom = 'Bfrtip', buttons = c('csv')))
  dt_options <- dt_options[-which(names(dt_options) == 'options')]
  dt_options <- c(dt_options, list(rownames = FALSE,
                                   filter = 'top',
                                   class = 'cell-border stripe',
                                   extensions = 'Buttons', 
                                   options = options_list))
  
  return(do.call(DT::datatable, 
                 c(list(data,
                        caption = caption), 
                   dt_options)))
}
```


```{r}
#assume data has already been aglomerated

fslong_data <- readRDS('sea_rsfc_fslong_schaefer400x7.RDS')
sesreg_data <- readRDS('sea_rsfc_schaefer400x7.RDS')
schaefer_400_7_net_ids <- readRDS('schaefer_ids.RDS')

fslong_data_wi <- get_within_net_connections(fslong_data)
sesreg_data_wi <- get_within_net_connections(sesreg_data)
```

The total amount of variance per network is higher in the FreeSurfer Longitudinal registration.

```{r}
#what's the total amount of variance per network?
var_data <- data.table::rbindlist(list(fslong = fslong_data_wi[, list(variance = var(z)), by = 'net1'],
                                       sesreg = sesreg_data_wi[, list(variance = var(z)), by = 'net1']), 
                                  idcol = 'reg')
var_data[, sd := variance^.5]
var_data$source <- 'total observed'
var_data_w <- dcast(var_data, net1 ~ reg)
setnames(var_data_w, c('net1', 'fslong', 'sesreg'), c('Network', 'FS Long', 'By Session'))
knitr::kable(var_data_w, caption = 'Total Variance', digits = 3)
```

What's the amount of variance explained by subject and session?

```{r}
grab_fits_var <- function(netname, registration){
  fslong_text <- ''
  if(registration == 'fslong'){
    fslong_text <- 'fslong-'
  }
  file_name <- file.path('models', paste0('lmer_fit-3l-', fslong_text, netname, '.RDS'))

  afit <- readRDS(file_name)
  afit_var <- get_variance_table(afit)
  
  comp_var_table <- afit_var[afit_var$out_of == 'total', c('source', 'variance', 'sd', 'percent')]
  comp_var_table <- as.data.table(rbind(comp_var_table, 
                                        c('Total', 
                                          sum(comp_var_table$variance), 
                                          sum(comp_var_table$variance)^.5, 
                                          100, 'Cont')))
  comp_var_table$net1 <- netname
  comp_var_table$reg <- registration
  return(comp_var_table)
}


models_list <- apply(expand.grid(netname = unique(fslong_data_wi$net1), 
                                 registration = c('fslong', 'sesreg')),
                     1, as.list)

tables <- lapply(models_list, do.call, what = grab_fits_var)

tables_dt <- rbindlist(tables)
```

The FreeSurfer longitudinal pipeline registrations (FS-long) yield more overall variance than the session-by-session registration (By Sess). However, the models using the FS-long registration data yield more variance explained by session and participant random effects, both in terms of percentage of total variance, as well as raw variance explained.

```{r}
format_percent <- function(p){
  sprintf('%.1f%%', p)
}
format_num <- function(r){
  sprintf('%.4f', r)
}
prop_resid_explained <- tables_dt[source %in% c('Total', 'Residual'), 
                                  c('net1', 'source', 'percent', 'variance', 'reg')]
prop_resid_explained_l <- data.table::melt(prop_resid_explained, id.vars = c('net1', 'source', 'reg'))
prop_resid_explained_l[, key := do.call(paste, c(.SD, sep = '_')), .SDcols = c('variable', 'source')]
prop_resid_explained_l[, c('source', 'variable') := NULL][, value := as.numeric(value)]
prop_resid_explained <- data.table::dcast(prop_resid_explained_l, ... ~ key, value.var = 'value')
prop_resid_explained[, c('perc_resid_expl', 'var_resid_expl') := 
                       list(percent_Total - percent_Residual, variance_Total - variance_Residual)]
prop_resid_explained[, c('percent_Residual', 'percent_Total') := NULL]
setnames(prop_resid_explained, c('variance_Residual', 'variance_Total'), c('var_resid', 'var_tot'))
prop_resid_explained_l <- data.table::melt(prop_resid_explained, id.vars = c('net1', 'reg'))
prop_resid_explained_l[, key := do.call(paste, c(.SD, sep = '-')), .SDcols = c('variable', 'reg')]
prop_resid_explained_w <- dcast(prop_resid_explained_l, net1 ~ key, value.var = 'value')
perc_cols <- grep('perc_', names(prop_resid_explained_w), value = TRUE)
numcols <- grep('(net1|perc_)', names(prop_resid_explained_w), value = TRUE, invert = TRUE)
prop_resid_explained_w[, (perc_cols) := lapply(.SD, format_percent), .SDcols = perc_cols]
prop_resid_explained_w[, (numcols) := lapply(.SD, format_num), .SDcols = numcols]
colnames <- 
  c('Network',#1
    'FS-long: % Variance Explained',#2
    'By Sess: % Variance Explained',#3
    'FS-long: Residual Variance',#4
    'By Sess: Residual Variance',#5
    'FS-long: Variance Explained',#6
    'By Sess: Variance Explained',#7
    'FS-long: Total Variance',#8
    'By Sess: Total Variance')#9
setnames(prop_resid_explained_w, colnames)
colnames <- colnames[c(1,8,9,2,3,6,7,4,5)]
setcolorder(prop_resid_explained_w, colnames)
knitr::kable(prop_resid_explained_w, caption = "Variance explained by session and ID random effects")
# dt_table(prop_resid_explained_w, caption = "Variance explained by session and ID random effects", 
#          dt_options = list(options = list(
#            columnDefs = list(list(className = 'dt-right', targets = 1:8)))))
```

```{r}

toplot <- prop_resid_explained_l[variable %in% c('var_resid', 'var_resid_expl')]
toplot[, variable := factor(variable, 
                            levels = c('var_resid_expl', 'var_resid'), 
                            labels = c('Random Effects', 'Residual'))]
ggplot(toplot,
       aes(x = reg, y = value, group = variable, fill = variable)) + 
  geom_col() + 
  facet_grid(. ~ net1) + 
  scale_fill_manual(values = apal[c(4,1)]) + 
  jftheme + 
  theme(axis.text.x = element_text(hjust = 0, angle = 360-45)) + 
  labs(y = 'Variance', x = 'Registration Strategy', fill = '')
```